---
title: "Computational Musicology"
author: "Bente Klein Hazebroek"
date: "25 February 2019"
output: 
    flexdashboard::flex_dashboard:
      orientation: columns
      vertical_layout: fill
      storyboard: true
---

```{r}
library(tidyverse)
library(spotifyr)
library(flexdashboard)
library(plotly)
library(Cairo)
library(shiny)
Sys.setenv(SPOTIFY_CLIENT_ID = '7695c57a82654e18ae7990e714b60c38')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'fac65dcc16154cbeb79e71ef5dbd9465')
```

# Computational Musicology
*Bente Klein Hazebroek*  **I**  *11863226*  **I**  *February 2019*
============================================================================================================

### My Corpus
------------------------------------------------------------------------------------------------------------
I am looking at Spotify's 'Coffeehouse' and 'On the Road' playlists. A coffeehouse is a pleasant place where you can talk, work or read while enjoying a cup of coffee; you will probably listen to easy, mellow songs here. But, when you're on the road, you have to kill time while going from A to B (whether you're alone or with friends; you go to work or are on a holiday). Since the car is a private place, you might listen to uplifting, sing-along songs. What are the characteristics of these playlists and in what way do they differ? I am using the following playlists for this research, composed by Spotify and Digster Nederland:

  **Coffeehouse:**
  
    1. Coffeehouse - Feel the vibe of the Coffeehouse playlist, full of relaxing singer-songwriter an pop music.    
    2. 't Koffiehuis - 'n Bakkie en gemoedelijke muziek op de achtergrond.    
    3. Your Favorite Coffeehouse - Curl up in your favorite spot with some sweet, mellow tunes...
  
 **On the Road:**
 
    1. Classic Road Trip Songs - The ultimate playlist to fuel your good mood while on the road.
    2. Onderweg - Met deze playlist heb je altijd de beste hits bij de hand voor onderweg.
    3. Road Trip - Go on a weekend getaway with your barkada and your favorite hits!
    
This results in a total of 281 Coffeehouse songs and 168 On the Road songs.

```{r}

```


```{r}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
